Return of the Linguist:\\Toward scalable parallel corpus generation for machine translation

### 2016-07-15 doc/paper.tex
I guess that's it for text - my GOODNESS the table in section 3.3 is sloppy and impenetrable...

@@

% website limits abstract to 300-word, but that's WAY too long. Mine is about 70 words, and it's short but not too short.
\begin{abstract}
The large-scale generation of synthetic parallel corpora for data-driven machine translation is proposed\amtaonly{ in this demo paper}.
A rudimentary system for doing this is presented, with mechanisms for syntactic complexity in the spirit of classical rule-based machine translation and high-quality sentence pairs in the spirit of example-based machine translation.
Preliminary experiments integrating these synthetic corpora into a shared task are reported.
Promising future directions are discussed, in particular, large-scale contributions from ``amateur linguists'' via crowdsourcing.
\end{abstract}

@@

The {\small \tt generator} module traverses the backbone of the ``interlingual'' tree from the {\small \tt nodes} module and recursively builds up parallel surface forms (bilingual sentence pairs).
Language-specific rules (e.g., word order) are implemented as ad hoc bits of code in this module,\footnote{
    Formally, the ``rule-based'' component of our system can be considered an extremely rudimentary synchronous (bilingual) context-free grammar of sorts \mycitep{transduction} that also admits long-range dependencies and semantic constraints.
    But it would be a disservice to those other well-developed methods to refer to our system as anything other than an ad hoc way to generate multilingual sentences.
%    MEH: Also, example-based mode admits the possibility of just writing down translations and then generating parse trees on PORTIONS of those templates.
%    Although then I guess you have a transductive grammar on those portions?
} reading long-range dependency information from the tree as needed.
All language-dependent code is confined to this module, although this is coupled with language-dependent information from the YAML data (e.g., the tags and punctuation in Figure~\ref{fig:yaml-rb}). 
% TODO: More ideally, A desirable longer-term goal would be to move language rules to data, but that's beyond the scope of this paper.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\noindent\makebox[\linewidth]{\rule{\paperwidth}{0.4pt}} % full-width horizontal line
TODO

@@

We briefly list
and how they are implemented in terms of the tree representation from Section~\ref{subsec:software}

a couple of reasons for keeping linguistic phenomena to the most basic ones (aside from manageability)
- we believe that for the purposes of generating parallel corpora, is most effective to treat very difficult, non-compositional phenomena (example?!) by simply writing them down, in tradition of example-based MT
- keeping the data format simpler makes it potentially easier to scale up to admit contributions from non-linguists

- emphasize that we prioritize generating in-domain sentences above linguistic soundness

% transformations were chosen with an eye toward scalability, so that things like semantic constraints (<animal> eats <food>) could be used (eating <food>, and should modify <animal>)

% if i'm gonna do source code snippets with zh, maybe it'd be best to use IMAGES? it's not for camera-ready anyway...
% well, i can do the meta-template easily... the problem is with REAL templates

@@

%Oh, maybe I should make this a table. that should cut down on the vertical spacing
%
%\begin{table}[ht]
%\centering
%\begin{tabular}{|l|l|}
%\hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%Gold en & 60,000. \\ \hline
%\end{tabular}
%\caption{ 
%    Linguistic constructs implemented in our multilingual generator.
%}
%\label{tab:implemented_linguistics}
%\end{table}

\begin{itemize}
\item clauses: transitive, meta, modal. a word on modalish. 
\item NP
\item ADJP
\item participles as transformations of clauses that convert subject into a linked node, allowing reuse of semantic constraints
\item prepositions as relations between two nodes (noun/noun or noun/verb), and accompanying semantic requirements on both object and target
\item custom nodes - because sometimes, the most effective way to represent a translation is to just write it down. (``template-based mode''): lots of iterals, maybe mark an NP somewhere

% other (if there's room)
\item ADVP (targeting ADJP or Clause), pronouns with antecedents 
\item indirect object as a special case of ``null preposition'' - not necessarily linguistically well-founded, but good enough to make sentences
\item past tense as transformation of present tense  % MEH
\item other transformations: topicalization, and some hokier stuff like \zh{把} 
\end{itemize}

% big question: do I want to show a source code example?? yeah, i probably have room
\noindent\makebox[\linewidth]{\rule{\paperwidth}{0.4pt}} % full-width horizontal line
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
} % amtaonly

@@

\subsection{Data Snooping}
\textbf{
Warning: This section is dedicated to ``data snooping'' experiments that included information from the test set during training in one way or another.
The scores given here must not considered indicative of actual out-of-sample system performance.
}

@@

We believe the most promising path to large-scale parallel corpus creation is through educational crowdsourcing.
As with the success story of reCAPTCHA \mycitep{recaptcha}, we believe this will scale best when there is a natural harmony between data supply and demand.
Unlike other staple NLP tasks like parsing or part-of-speech tagging, MT is particularly well-suited to this, with English being taught as a foreign language in classrooms around the world.
The data format (English itself) is widely agreed upon, and there is even an existing market, a vast untapped resource of student problem sets that is yet to be harnessed by machine learning.
\amtaonly{ % Uh, should I bother with this stupidity, 
%As a back of the envelope estimate, at a translation rate of about 1 noun per minute (based on our own data input experience), even for a modest user base of 10,000 users, a ten-minute problem set could in principle yield 100,000 data samples.
%It seems like the bottleneck would be judicious problem set design.
It is also worth noting that in a sense, this could regarded as a ``massively-online'' version of elicitation-based techniques  from fieldwork linguistics \mycitep{elicitation}, with an emphasis on acquiring unprocessed parallel sentences, and doing so at scale.
}

>>>>

    \usepackage{tabularx}

@@

% website limits abstract to 300-word, but that's WAY too long. Mine is about 70 words, and it's short but not too short.
\begin{abstract}
The large-scale generation of synthetic parallel corpora for data-driven machine translation is proposed\amtaonly{ in this demo paper}.
A rudimentary system for doing this is presented, with mechanisms for syntactic complexity in the spirit of classical rule-based machine translation and high-quality sentence pairs in the spirit of example-based machine translation.
Preliminary experiments integrating these synthetic corpora into the IWSLT shared task are reported.
In the process, it was incidentally found that correcting $\sim$4\% of the shared task's training corpus for obvious, clerical-type errors resulted in a $\sim$6\% relative increase in BLEU scores.
Promising future directions are discussed, in particular, large-scale contributions from ``amateur linguists'' via crowdsourcing.
\end{abstract}

@@

The {\small \tt generator} module traverses the backbone of the ``interlingual'' tree from the {\small \tt nodes} module and recursively builds up parallel surface forms (bilingual sentence pairs).
Language-specific rules (e.g., word order) are implemented as ad hoc bits of code in this module,\footnote{
    Formally, the ``rule-based'' component of our system can be considered an extremely rudimentary synchronous (bilingual) context-free grammar of sorts \mycitep{transduction} that also admits long-range dependencies and semantic constraints.
    But it would be a disservice to those other well-developed methods to refer to our system as anything other than a makeshift way to generate multilingual sentences.
%    MEH: Also, example-based mode admits the possibility of just writing down translations and then generating parse trees on PORTIONS of those templates.
%    Although then I guess you have a transductive grammar on those portions?
} reading long-range dependency information from the tree as needed.
All language-dependent code is confined to this module, although this is coupled with language-dependent information from the YAML data (e.g., the tags and punctuation in Figure~\ref{fig:yaml-rb}). 
% TODO: More ideally, A desirable longer-term goal would be to move language rules to data, but that's beyond the scope of this paper.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%TODO

@@

Finally, to conclude this section, we briefly list linguistic constructs and their implementations in terms of the tree representation from Section~\ref{subsec:software}.
Note that since we are interested primarily in the generated surface forms, the linguistic soundness of how they are implemented in our tree representation is of relatively secondary importance.
Aside from manageability, there are a few other reasons for focusing on just the most basic linguistic phenomena for the ``rule-based'' mode.
We strongly believe that for the purposes of generating large and high-quality parallel corpora, it is most effective to treat difficult, non-compositional phenomena by simply writing them down, in tradition of example-based MT.
Also, minimizing the learning curve makes it potentially easier to scale up to admit contributions from non-linguists.

% transformations - with scalability in mind, so that things like semantic constraints (<animal> eats <food>) could be used (eating <food>, and should modify <animal>)

@@

%Oh, maybe I should make this a table. that should cut down on the vertical spacing
%
\begin{table}[ht]
\centering
%\begin{tabular}{|l|l|} % tabularx allows automatic wrapping
\begin{tabularx}{\linewidth}{ |l|X| }
\hline
Clauses & indicative present active; intransitive, transitive, ``meta'' (clause subnode), ``modalish'' (e.g., ``Alice wants to kick Bob'' - with ``Alice'' linked to ``kick'' as in Figure~\ref{fig:tree} for subject-verb semantics) \\ \hline % see with a ``dependency link'' for subject-verb syntactic agreement and 
NP & noun phrases headed by nouns, names, or pronouns; optional modifiers as child nodes; pronoun antecedents via links \\ \hline
ADJP & adjectives as NP modifiers, with links to head nouns (e.g., this vs. these) \\ \hline
ADVP & adverbs as Clause or ADJP modifiers \\ \hline
Participles & transformations of clauses that convert subject child node into a linked node, allowing reuse of semantic constraints \\ \hline
PP & prepositional phrases (relations) between two nodes (noun/noun or noun/verb), with relation-dependent semantic requirements on both nodes; indirect objects as ``null preposition'' nodes \\ \hline
%Custom nodes & ``example-based'' nodes as discussed in Section~\ref{subsec:software} \\ \hline
%Transformations & alterations of existing templates - past tense, topicalization, \zh{把}-construction \\ \hline
\end{tabularx}
\caption{ 
    Linguistic constructs implemented for the ``rule-based'' mode of our generator.
}
\label{tab:implemented_linguistics}
\end{table}

%\begin{itemize}
%\item 
%\item participles as

%
%% other (if there's room)
%\item indirect object as a special case of ``null preposition'' - not necessarily linguistically well-founded, but good enough to make sentences

%\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
} % amtaonly

@@

\subsection{Data Snooping}
\textbf{
Warning: This section is dedicated to ``data snooping'' experiments that included information from the test set during training in one way or another.
The scores given here must not be considered indicative of actual out-of-sample system performance.
}

@@

We believe the most promising path to large-scale parallel corpus creation is through educational crowdsourcing.
As with the success story of reCAPTCHA \mycitep{recaptcha}, we believe this will scale best when there is a natural harmony between data supply and demand.
Unlike other staple NLP tasks like parsing or part-of-speech tagging, MT is particularly well-suited to this, with English being taught as a foreign language in classrooms around the world.
The data format (English itself) is widely agreed upon, and there is even an existing market, a vast untapped resource of student problem sets that is yet to be harnessed by machine learning.
\amtaonly{ % Uh, should I bother with this stupidity, 
%As a back of the envelope estimate, at a translation rate of about 1 noun per minute (based on our own data input experience), even for a modest user base of 10,000 users, a ten-minute problem set could in principle yield 100,000 data samples.
%It seems like the bottleneck would be judicious problem set design.
It is also worth noting that in a sense, this could regarded as a ``massively-online'' version of elicitation-based techniques from fieldwork linguistics \mycitep{elicitation} for low-resource language pairs, with an emphasis on acquiring unprocessed parallel sentences, and doing so at scale.
}
